---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
read data:

```{r}
dat <- read.csv('./data/activity.csv')

```

remove missing values (NA) from data 

```{r}
bad <- is.na(dat$steps)

sum_bad <- sum(bad)

xdat <- dat[!bad,]
```
There are `r sum_bad` missing values. These will be ignored for the following analysis.

## What is mean total number of steps taken per day?

Group data by-day with dplyr package
```{r}
library(dplyr)
xdat_day <- group_by(xdat, date)
```

Sum steps for each day
```{r}
xdat_day_sum <- summarize(xdat_day, sum = sum(steps), mean = mean(steps), median = median(steps))

```

Plot histogram of steps per day with ggplot
```{r section,  tidy=TRUE, messages=FALSE}
library(ggplot2)
qplot(sum, data = xdat_day_sum, geom = "bar")
```

Table summary of mean/median number of steps

```{r showtable, results = "asis"}
library(xtable)
xt <- xtable(summary(xdat_day_sum))
print(xt, type = 'html')
```

## What is the average daily activity pattern?

First group data by-interval

```{r}
xdat_interval <- group_by(xdat, interval)
```

Calculate average steps of daily intervals

```{r}
xdat_interval_mean <- summarize(xdat_interval, mean = mean(steps))
```

Convert intervals to continuous time. [Reference] (https://class.coursera.org/repdata-012/forum/thread?thread_id=36)

```{r}
interval <- xdat_interval_mean$interval
intToHHMM <- function (i) {  sprintf("%02d:%02d", i %/% 100, i %% 100)}
p <- intToHHMM(interval)
q <- strptime(p, format = "%H:%M")
```

Add converted time to data

```{r}
xdat_interval_mean <- cbind(xdat_interval_mean, q)
```

Plot:

```{r}
qplot(q, mean, data = xdat_interval_mean, geom = "line" , xlab = "5-minute interval")
```

Peak time of maximum number of steps is 8:35AM

## Imputing missing values

There are `r sum_bad` missing values. 

The average daily steps of 5-min intervals will be added for this analysis.

Obtain rows of data with NA

```{r}
baddat <- dat[bad,]
```

Vectorize average daily steps by 5-min interval

```{r}
avedat <- xdat_interval_mean$mean
```

Assign "NA" with average daily steps

```{r}
baddat$steps <- avedat
```

Combine previous "cleaned" data with assigned data to NAs

```{r}
alldat <- rbind(xdat, baddat)
```

Sort by date

```{r}
new_dat <- arrange(alldat, date)
```

Group data by-day

```{r}
new_dat_day <- group_by(new_dat, date)
```

Sum steps for each day

```{r}
new_dat_day_sum <- summarize(new_dat_day, sum = sum(steps), 
                             mean = mean(steps), median = median(steps))
```

Generate new data-frame w/ and w/out NAs

```{r}
df <- rbind(data.frame(NAs = "not_included" , obs = xdat_day_sum$sum), 
            data.frame(NAs = "included" , obs = new_dat_day_sum$sum))
```

Histogram of combined data:

```{r}
qplot(obs, data = df, 
      geom = "bar", position = "dodge", fill = NAs, xlab = "# of steps")
```

Table summary of mean/median/sum for entire data and data w/o NA below

```{r showtable2, results = "asis"}
library(xtable)
xt2 <- xtable(summary(new_dat_day_sum))
print(xt2, type = 'html')
print (xt, type = 'html')
```

The means don't change based on this analysis... 

## Are there differences in activity patterns between weekdays and weekends?

convert dates to weekdays

```{r}
dates <- new_dat$date
dates_POS <- as.POSIXlt(dates)
dates_POS_day <- weekdays(dates_POS)
```

Vectorize days into 'weekday' or 'weekend'

```{r}
intoDay <- function (a) {
        day <- c()
        for (i in a){
                {if (i == "Saturday" | i == "Sunday") {
                        day <- c(day, "weekend")}

                else {
                        day <- c(day, "weekday")}
                }
        }
        return(day)
}

day <- intoDay(dates_POS_day)
```

Bind day category vector to data (with NAs included)

```{r}
new_dat <- cbind(new_dat, day)
```

Group by-day and by-interval

```{r}
new_dat_interval_day <- group_by(new_dat, interval, day)
```

Obtain averages of steps for weekend and weekdays

```{r}
new_dat_interval_day_sum <- summarize(new_dat_interval_day, mean_steps = mean(steps))
```

Separate weekday and weekend data

```{r}
new_dat_interval_day_sum_weekday <- filter(new_dat_interval_day_sum, day == "weekday")
new_dat_interval_day_sum_weekend <- filter(new_dat_interval_day_sum, day == "weekend")
```

Add contiuous time 

```{r}
new_dat_interval_day_sum_weekday_q <- cbind(new_dat_interval_day_sum_weekday, q)
new_dat_interval_day_sum_weekend_q <- cbind(new_dat_interval_day_sum_weekend, q)
```

Make one data frame from weekend & weekday mean steps

```{r}
df <- cbind(data.frame(interval = new_dat_interval_day_sum_weekday_q$q, weekday_steps = new_dat_interval_day_sum_weekday_q$mean_steps, weekend_steps = new_dat_interval_day_sum_weekend_q$mean_steps))
```

plot weekday and weekend pannels:

```{r}
par(mfcol = c(2,1))
with(df,{
        plot(df$interval, df$weekday_steps, type = 'l', xlab = "time" , ylab = "average steps" , main = "weekday")	
	plot(df$interval, df$weekend_steps, type = 'l', xlab = "" , ylab = "average steps" , main = "weekend")
})
```






